---
title: "Creating an alluvial plot with R"
author:
  - name: Lisa Hefele
    url: {}
date: 2024-10-05
categories: ["Data visualization"]
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```






## Step 1: Laod necessary packages


```{r echo=TRUE}

library(ggplot2)
library(ggalluvial)
library(dplyr)

```





## Step 2: Create a fictional dataset

This dataset now reflects a situation where participants from lower wealth and less educated backgrounds are less likely to be vaccinated. It is completely fictional.


```{r create example data, echo=TRUE}
set.seed(123)  # For reproducibility

# Simulate a dataset
n <- 1000  # Number of participants

# Wealth categories (lower wealth indices are less likely to be vaccinated)
wealth <- factor(sample(c("Low", "Medium", "High"), n, replace = TRUE, prob = c(0.5, 0.3, 0.2)))

# Mother's education levels (less educated mothers are less likely to vaccinate)
education <- factor(sample(c("No Education", "Primary", "Secondary"), n, replace = TRUE, 
                           prob = c(0.5, 0.3, 0.2)))

# Region is randomly assigned for now
region <- factor(sample(c("North", "South", "Central"), n, replace = TRUE))

# Antenatal care visits (a binary factor)
antenatal_care <- factor(sample(c("Yes", "No"), n, replace = TRUE))

# Define a probability of vaccination based on wealth and education
vaccination_prob <- ifelse(wealth == "Low" & education == "No Education", 0.2,  # Low wealth and no education
                   ifelse(wealth == "Low" & education == "Primary", 0.3,       # Low wealth and primary education
                   ifelse(wealth == "Medium" & education == "No Education", 0.4, # Medium wealth and no education
                   ifelse(wealth == "Medium" & education == "Primary", 0.6,    # Medium wealth and primary education
                   ifelse(wealth == "High" & education == "No Education", 0.6, # High wealth and no education
                   0.8)))))  # Higher probability for other combinations

# Simulate vaccination status based on probability
vaccinated <- factor(rbinom(n, 1, vaccination_prob), labels = c("No", "Yes"))

# Combine into a data frame
df <- data.frame(wealth, education, region, antenatal_care, vaccinated)

# View the first few rows of the dataset
head(df)


```



## Step 3: reshape the data 

```{r reshape data, echo=TRUE}
# Group by the demographic variables and count occurrences
df_alluvial <- df %>%
  count(wealth, education, region, antenatal_care, vaccinated)

# View the data to make sure it's correct
head(df_alluvial)

```




## Step 4: Create an alluvial plot

In an alluvial plot, each "stratum" represents a category of a variable, and the "flows" represent how individuals move between these categories. Here is an example of how to make the plot:

```{r create alluvial plot, echo=TRUE}

# Create the alluvial plot
df_alluvial %>% 
  ggplot(aes(axis1 = wealth, axis2 = education, axis3 = region, axis4 = antenatal_care, y = n)) +
  geom_alluvium(aes(fill = vaccinated), width = 0.2) +  # Moved 'fill' inside aes()
  geom_stratum(width = 0.2, fill = "lightgrey", color = "black") +
  geom_text(stat = "stratum", aes(label = after_stat(stratum))) +
  scale_x_discrete(limits = c("Wealth", "Education", "Region", "Antenatal Care"), expand = c(0.15, 0.05)) +
  labs(title = "Alluvial diagram of vaccination status",
       y = "Number of participants",
       x = "Demographic variables") +
  theme_minimal()



```


















