---
title: "High exposure to hepatitis C virus in Saravan, southern Laos: Identification of several risk practices"

author:
  - name: Lisa Hefele
    url: 
date: 2024-05-18
categories: ["Research", "Publications"]
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



## Hepatitis C Virus risk factors study in Southern Laos

In a follow-up study conducted in Samuoi district, Saravan province, southern Laos, we investigated the prevalence and risk factors of hepatitis C virus (HCV) infection. Our findings revealed a high seropositivity rate, with 39.8% of participants testing positive for HCV antibodies. Notably, non-sterile medical practices, including the use of re-used syringes (72.5%) and unlicensed parenteral treatments (58%), were common. Tattoos and accidental blood exposure were significantly associated with HCV infection. This study underscores the urgent need for community-wide HCV testing and awareness campaigns to reduce transmission risks. Further viral sequencing is ongoing to enhance our understanding of HCV epidemiology in the region.

For more details, you can access the full paper [here](https://www.thelancet.com/journals/lanwpc/article/PIIS2666-6065(22)00247-4/fulltext).




```{r prep map, echo=FALSE, message=FALSE, warning=FALSE, results='hide', out.width="100%", fig.cap="Map highlighting the district of Samuoi."}

# Load necessary libraries
library(sf)
library(tmap)

# Read shapefiles (adjust the file paths)
country <- suppressWarnings(st_read("lao_adm_ngd_20191112_SHP/lao_admbnda_adm0_ngd_20191112.shp"))
province <- suppressWarnings(st_read("lao_adm_ngd_20191112_SHP/lao_admbnda_adm1_ngd_20191112.shp"))
districts <- suppressWarnings(st_read("lao_adm_ngd_20191112_SHP/lao_admbnda_adm2_ngd_20191112.shp"))

# Filter the district you want to highlight
highlighted_district <- districts[districts$ADM2_EN == "Samuoi", ]


```

```{r map, echo=FALSE, message=FALSE, warning=FALSE, out.width="100%", fig.cap="Map highlighting the district of Samuoi."}

# Create a tmap object
map <- tm_shape(country) + 
  tm_borders() + 
  tm_shape(province) + 
  tm_borders(lwd = 1.5, col = "cornflowerblue") +  # Province borders
  tm_shape(districts) + 
  tm_borders(lwd = 0.5, col = "azure4") +  # District borders
  tm_shape(highlighted_district) + 
  tm_fill(col = "darkmagenta", alpha = 0.7) +   # Highlight specific district
  tm_borders(lwd = 2, col = "darkorchid4") +  # Border around the highlighted district
  tm_layout(legend.outside = TRUE)

# View the map
tmap_mode("view")  # Interactive mode
map


```









